A Transdiciplinary approach to developing a 3D spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities  around the coast of South Africa

### 2019-10-23 literature_review.Rmd
additions and spelling, grammer check

Marine ecosystems are maintained by a variety of complex interactions between abiotic and biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology (Doney et al. 2011; Harley et al. 2006; Burrows et al. 2011; McGowan et al. 1998). These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems (Guimaraes & Coutinho 1996; Harley et al. 2006; Poloczanska et al. 2013; Jennings & Brander 2010; Polovina 2005; Johnson et al. 2011; Wernberg et al. 2016; Dayton et al. 1999). Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and overfishing (Blamey et al. 2012; Blamey et al. 2015). Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean (Guimaraes & Coutinho 1996; McGowan et al. 1998; McQuaid & Branch 1984; Laufkötter et al. 2015; Belkin 2009; Filbee-Dexter et al. 2018; Miller et al. 2011; Smale et al. 2011; Smale & Moore 2017). In order to persist and survive within variable and changing environments, organisms must either migrate, adapt, or die.

@@

Pollution is a huge environmental problem that affects both terrestrial and marine ecosystems. Pollution from land enters the sea where it could harm or kill marine organisms or is transported by ocean currents to other coastal areas. In these areas the pollution could re-enter the ocean or be blown by wind into terrestrial ecosystem where it could once again be ingested by organisms causing harm or even death. Pollution is therefore a major threat to the environment and marine organisms. In recent years the effect of microplastics on the ocean has gained much traction with scientists and politicians alike. Microplastics cosiste of tiny particles of plastic and other pollution and are therefore are difficult to detect. Currently there is no hydrodynamic model that is able to determine dispersion, source and accumulation of microplastics along the South African coastline. Given the significant detrimental effect microplastics play in the ocean, it is important that such a mechanism be developed that will aid in better management of marine pollution in South Africa. Furthermore, this project allows for a multidisciplinary approach to be taken by combining ecology and coastal oceanography.

@@

Advances in numerical modelling has gained much traction in recent years and has been applied in a variety of ways with regards to ecological studies. For example, a study by Wang and Xia (2009) used the Delft3D-Flow model to assess the hydraulic suitability of a stream as a spawning ground for the Chinese Sturgeon (Acipenser sinensis) in the Yangtze River. The authors calculated the horizontal mean vorticity which was used to assess the hydraulic environment of spawning ground. The flow field state was determined through model simulation and field-measured data used to validate the model. The results added to existing scientific database for spawning ground hydraulic environmental protection. Different numerical models can often be integrated to model across ecosystem levels. For example a study by Leon et al. (2003) used integrated physical (Delft3D hydrological model) and bio-chemical (Agricultural Non-point Source model) processes models to investigate the possible impact on the Lake Malawi water quality due to management actions performed at the watershed level.

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are no exception. Wave exposure has been shown to play a role in determining distribution, abundance, diversity, composition, growth (Cousens 1982) and productivity (Pedersen and Nejrup 2012) of macroalgae communities. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communities indirectly through the alteration in effect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundary layer around the frond (???). The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgement, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarily in the marine environment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macroalgae have been able to persist in often harsh and variable wave environments. Macroalgae are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed genetic traits. A study by Koehl et al. (2008) showed that transplanted Nereocystis luetkeana plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Advances in genetic techniques and taxonomy have revealed that species delineation based on morphology has been inaccurate, and organisms that were once considered two separate species are actually one species. For example, Moss (1948) investigated the anatomy, chemical composition of Fucus spiralis at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag. Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macroalgae. Duggins et al. (2003) examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included Costaria costata, Agarum fimbriatum, and Laminaria complanata and *Nereocystis luetkeana*. The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

In the wave swept nearshore one would expect organisms to reflect the harsh hyrdrodyanmic environment by being streamlined, small and amoured. This is certianly the case for an array of fauna which often comprise of hard, rigid bodies that are held firmly in place to the substratum such as limpets and isopods.

@@

The generating force and restoring force occur along these boundaries, the generating force pushes water up across ocean boundaries while the restoring force pulls it back to where the boundary was originally, trying to restore the balance in energy. This "tug of war" between the two forces creates an oscillating motion between the boundary layers which acts as a point of disturbance, sending out energy in all directions. The disturbance energy will continue to manifest itself provided the tug and pull between the two forces is occuring. Once the generating force stops, and ultimately the energy from the point of disturbance, the waves dissipate and water restores to its original state. A simple example would be blowing on water in a cup. The blowing of air onto the surface of the water in the cup creates a point of disturbance that pushes the air boundary into the water boundary. The restoring force is the surface tension of the water that is maintained through hydrogen bonds between water molecules. Once blowing on the water in the cup has stopped, the ripples or “waves” in the cup begin to diminish in size until the surface tension returns to its original state. The hydrogen bonds between the water molecules are stronger than the force of gravity, and therefore the force of the surface tension returns the water to its original state. These waves are known as “capillary waves” and is essentially residual energy after the generating source has stopped. The restoring force can also be in the form of gravity and are known as “gravity waves”. Using the same example, if one blows too hard, the water boundary is pushed up into the air boundary, breaking the hydrogen bonds between water molecules which allows gravity to return the water to its original state. All waves in the ocean are either capillary waves or gravity waves, and their classification will be dependent on the restoring force involved.

@@

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obvisiously not he case. This is known as *surface elevation* and is the instaneous elevation of the sea surface abouve a specific point in a time record (see figure \ref{fig:surface elevation figures}). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* or *upwardward zero-crossings* (see figure \ref{fig:surface elevation figures}). Both zero-crossings are symmetrical and essentially the same statistically. However, in practice the downward zero-crossings are preferred as it takes steepness of a approaching wave into account (the front, see figure \ref{fig:surface elevation figures}) which is relevant to characterising breaking waves. It is important to note that surface elevation can be negative while a wave profile cannot.

@@

Waves are complex and various approaches, techniques and devices have been developed in order to measure waves effectively. One of the ways that has been used extensively used in the past but less so today, are visual estimates which can be used to characterise *significant swell height* (Hs) and *significant swell period* (Tp). Visual estimates use 15 - 20 of the most well defined, higher waves of a number of wave groups to characterise Hs and Tp. Although these parameters are useful, they do not accurately reflect the waves that are occurring in nature. Ocean waves are a combination of wind sea (short, irregular, locally generated waves) and swell (long, smooth waves, generated by distant storms) and so more parameters are needed to separate Hs and Tp driven by wind sea and swell, i.e. separate Hs and Tp parameters for wind sea and swell. However, even with separate parameters for the different types of waves would still not be enough to effectively characterise waves in the ocean. In order to characterise the detail and complexity of ocean waves a different approach must be used, know as the *spectral* approach. This approach is based on the idea that the sea surface can be characterised as the summation of a large number of harmonic wave components, calculated by the *random-phase/amplitude* model. The random-phase amplitude model is a summation of wave components on a discrete scale, however a continous scale is more relevant to characterising and measuring of waves in the ocean. The discrete spectrums are converted to continoius spectrums through various arithmetic methods. Therefore, the random-phase amplitude model is used to obtain a *continious variance density spectrum* which is a statistical relevant measure for scientists and engineers.

@@

An object will only float if the object density ($\rho$) is less than or equal to the density of the liquid in which the object is submerged, i.e. $\rho_{object} < \rho_{liquid}$. The bouyancy force (*BF*) is the force of the displaced liquid pushing the object in an upwards direction which is equal to the weight of the displaced liquid and is expressed as: $$BF = W_{displaced-liquid}$$ The weight of the displaced liquid can be calculated by multiplying the mass of the liquid by the gravitational constant $g$ and so the weight of the displaced liquid can be calculated by: $$W_{displaced-liquid} = mg$$ The equation is expanded upon by including the calculation for mass, $m = \rho V$, and so now the weight of the displaced liquid can be calculated as: $$mg = \rho_{liquid} V g$$ where $\rho$ represents the density of the liquid and $V$ represents the portion of the object submerged. Since the object is floating, the BF can also be expressed by the weight of the object and as mentioned previously the weight of the displaced liquid can also be expressed by $mg$, which in turn can be written as $W_{displaced-liquid} = W_{object}$. Now through substituting the calculations for weight, volume and the dimensions of the object, the equation can be simplied to: $$x= h\displaystyle\frac{\rho_o}{\rho_l}$$ where $h$ is the overall height of the object, $\rho_o$ represents the density of the object and $\rho_l$ represents the density of the liquid.

It is important to note that this equation is only applicable to objects that are floating i.e. $\rho_{object} < \rho_{liquid}$. The portion of a floating object above the water surface can also be calcuated using a similar approach, and instead of calculating the submerged portion ($x$) the calculation will now solve for the fraction above the surface which is represented by $y$ (see figure \ref{fig:BF_1}). The same equation for calculating density ($\rho = \displaystyle\frac{m}{V}$) will also be used in solving for $y$. As mentioned previously, for floating objects the $BF = m_{object}g$ and the BF can also be defined as the weight of the displaced liquid ($BF = W_{displaced-liquid}$). Therefore, the weight of the displaced liquid is equal to the weight of the object ($W_{displaced-liquid} = W_{object}$). On this basis equation ?? can be written as $$m_{liquid}g = m_{object}g$$ Substituting the equation for density into equation ?? the equation is now: $$\rho_{o}V_{o}g = \rho_{l}V_{l}g$$ Important to note is that $V_{o}$ represents the volume of displaced liquid i.e not the entire object volume (see figure \ref{fig:BF_1}). The constant $g$ is removed through cancellation and now the equation can be written as: $$\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$$ We also now that the volume of the object submerged is equal to the volume of the displaced liquid and the fraction $\displaystyle\frac{V_l}{V_o}$ represents the portion of the object submerged. Therefore to calculate the portion above the surface this must be subtracted by 1, therefore: $$1-\displaystyle\frac{V_l}{V_o} = y$$ and since $\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$ the equation can now be solved via density of the liquid and the object: $$y = 1 - \displaystyle\frac{\rho_o}{\rho_l}$$ Solving the equation algebraically it can now be written as:$$ y = \displaystyle\frac{\rho_l - \rho_o}{\rho_l}$$

Once the fractions of submerged and non-submerged portions of the object have been calculated the next calculation relevant to this study would be to calculate the weight needed to cause the object to sink or as least be completely submerged. This will relate to epiphyte loading on kelp rafts in order to determine raft longevity. If a weight is placed on the object in the conceptual model that has enough weight to fully submerge (but not sink) the object then an addition bouyancy force is needed, which will be represented by $BF_2$ which is equal to the additional mass that has been added which can be expressed as $mg$. Based on the previous calculations and the laws of physics we know that: $$W_{object} = mg = BF_2 = W_{add-liq-displaced}$$ The weight of the additional liquid displaced can now be calculated by substituting the density and volume of the additional displaced liquid along with the gravitational constant $g$ which is expressed as: $$W_{AL} = \rho_{l}V_{AL}g$$
where the subscript $AL$ represents the components of *additional liquid displaced*. The $V$ can also represent the change in volume ($\Delta V$) and so the expression can be expanded to: $$W_{AL} = \rho(\Delta V) g$$} where $\Delta V$ represents the change in volume of the object, i.e $y$. To estimate the amount of input into the model the authors used a scaled approach to define the release of particles, and where based on previous research performed by @halpern2008.

@@

@@

$$u'= u + \frac{\delta tu_t}{2}$$ $$\upsilon'=\upsilon + \frac{\delta t\upsilon_t}{2}$$
and $u$, $\upsilon$ are orthogonal velocity components, $t$ is time, $\delta t$ is the model time step, $u_\chi$, $u_y$, $\upsilon_y$ are the $u$ and $\upsilon$ spatial velocity gradients, and $u_t$, $\upsilon_t$ are temporal gradients. Horizontal diffusion was modelled as a random walk with seperate longitudinal and lateral coefficients set to simulate random turbulence. The distance increments moved by the particle at each time step is calculated as:
$$\Delta X = R_N \sqrt{6E_1\delta t}$$ $$\Delta y = R_N\sqrt{6E_2\delta t}$$
where $R_N$ is a random number in the uniform range (-1,1) and $E_1,E_2$ are the longitudinal and lateral eddy diffusivities, respectivley.

@@

A study by @lebreton2012 proposed a methododology to track floating debris from source to sink based on discriptions of global waste production and ocean surface currents, using a combination of hydrodynamic and particle dispersion modelling. The particle tracking model applied uses a two stage process, (1) a hydrodynamic model solves the equations of motion to characterise the movement of water within the domain of the model, (2) virtual particles are released into the flow field established in the first step, and particles are allowed to move through hydrodynamic forcing. The authors extracted sea surface currents from the  HYCOM/NCODA ocean circulation modelling system. The HYCOM model is forced by the US Navy's Operational Global Atmospheric Prediction System (NOGAPS). The model also takes into account wind stress, wind speed, heat flux, and precipitation; and was looped five times to accuratley represent ocean circulation patterns for 30 years. The authors used the velocity data aquired from HYCOM and coupled it to the Lagrangian particle tracking model Pol3DD which was used to simulate dispersion. No additional wind stress terms were applied to the motion of particles, as this factor was already expressed in the HYCOM hydrodynamic data. Potential extra stress for emerged parts of debris was considered negligible as all floating debris are assumed to be fully submerged. The Pol3DD model was also modified so provide additional information on individual particles such as origin, age, and trajectory information of individual particles. Since the model considered a spherical earth, particles were allowed to re-enter the simulation if the passed around the 'globe'. In this study only the surface layer was considered as the prinicple layer which determines the fate of floating particles i.e anthorpogenic debris.

>>>>

Marine ecosystems are maintained by a variety of complex interactions between abiotic and biotic variables such as temperature, wave exposure, pH, competition, and processes such as top-down and bottom-up control, predator-prey relationships and phenology (Doney et al. 2011; Harley et al. 2006; Burrows et al. 2011; McGowan et al. 1998). These abiotic and biotic variables, the interactions between them, and the various ecological processes, ultimately determine the community composition and ecological functioning of all ecosystems (Guimaraes & Coutinho 1996; Harley et al. 2006; Poloczanska et al. 2013; Jennings & Brander 2010; Polovina 2005; Johnson et al. 2011; Wernberg et al. 2016; Dayton et al. 1999). Climate directly and indirectly affects the way in which abiotic and biotic variables interact, but is often compounded by other impacts such as habitat destruction, pollution, and over-fishing (Blamey et al. 2012; Blamey et al. 2015). Temperature and wave exposure have been recognised as important variables with regards to climate-driven changes within the ocean (Guimaraes & Coutinho 1996; McGowan et al. 1998; McQuaid & Branch 1984; Laufkötter et al. 2015; Belkin 2009; Filbee-Dexter et al. 2018; Miller et al. 2011; Smale et al. 2011; Smale & Moore 2017). In order to persist and survive within variable and changing environments, organisms must either migrate, adapt, or die.

@@

Pollution is a huge environmental problem that affects both terrestrial and marine ecosystems. Pollution from land enters the sea where it could harm or kill marine organisms or is transported by ocean currents to other coastal areas. In these areas the pollution could re-enter the ocean or be blown by wind into terrestrial ecosystem where it could once again be ingested by organisms causing harm or even death. Pollution is therefore a major threat to the environment and marine organisms. In recent years the effect of microplastics on the ocean has gained much traction with scientists and politicians alike. Microplastics consist of tiny particles of plastic and other pollution and are therefore are difficult to detect. Currently there is no hydrodynamic model that is able to determine dispersion, source and accumulation of microplastics along the South African coastline. Given the significant detrimental effect microplastics play in the ocean, it is important that such a mechanism be developed that will aid in better management of marine pollution in South Africa. Furthermore, this project allows for a multidisciplinary approach to be taken by combining ecology and coastal oceanography.

@@

Advances in numerical modelling has gained much traction in recent years and has been applied in a variety of ways with regards to ecological studies. For example, a study by Wang and Xia (2009) used the Delft3D-Flow model to assess the hydraulic suitability of a stream as a spawning ground for the Chinese Sturgeon (Acipenser sinensis) in the Yangtze River. The authors calculated the horizontal mean velocity which was used to assess the hydraulic environment of spawning ground. The flow field state was determined through model simulation and field-measured data used to validate the model. The results added to existing scientific database for spawning ground hydraulic environmental protection. Different numerical models can often be integrated to model across ecosystem levels. For example a study by Leon et al. (2003) used integrated physical (Delft3D hydrological model) and bio-chemical (Agricultural Non-point Source model) processes models to investigate the possible impact on the Lake Malawi water quality due to management actions performed at the watershed level.

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are no exception. Wave exposure has been shown to play a role in determining distribution, abundance, diversity, composition, growth (Cousens 1982) and productivity (Pedersen and Nejrup 2012) of macroalgae communities. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communities indirectly through the alteration in effect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundary layer around the frond (???). The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgement, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarily in the marine environment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macroalgae have been able to persist in often harsh and variable wave environments. Macroalgae are sessile organisms and incapable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed genetic traits. A study by Koehl et al. (2008) showed that transplanted Nereocystis luetkeana plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Advances in genetic techniques and taxonomy have revealed that species delineation based on morphology has been inaccurate, and organisms that were once considered two separate species are actually one species. For example, Moss (1948) investigated the anatomy, chemical composition of Fucus spiralis at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag. Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macroalgae. Duggins et al. (2003) examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included Costaria costata, Agarum fimbriatum, and Laminaria complanata and *Nereocystis luetkeana*. The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

In the wave swept nearshore one would expect organisms to reflect the harsh hydrodynamic environment by being streamlined, small and amoured. This is certainly the case for an array of fauna which often comprise of hard, rigid bodies that are held firmly in place to the substratum such as limpets and isopods.

@@

The generating force and restoring force occur along these boundaries, the generating force pushes water up across ocean boundaries while the restoring force pulls it back to where the boundary was originally, trying to restore the balance in energy. This "tug of war" between the two forces creates an oscillating motion between the boundary layers which acts as a point of disturbance, sending out energy in all directions. The disturbance energy will continue to manifest itself provided the tug and pull between the two forces is occurring. Once the generating force stops, and ultimately the energy from the point of disturbance, the waves dissipate and water restores to its original state. A simple example would be blowing on water in a cup. The blowing of air onto the surface of the water in the cup creates a point of disturbance that pushes the air boundary into the water boundary. The restoring force is the surface tension of the water that is maintained through hydrogen bonds between water molecules. Once blowing on the water in the cup has stopped, the ripples or “waves” in the cup begin to diminish in size until the surface tension returns to its original state. The hydrogen bonds between the water molecules are stronger than the force of gravity, and therefore the force of the surface tension returns the water to its original state. These waves are known as “capillary waves” and is essentially residual energy after the generating source has stopped. The restoring force can also be in the form of gravity and are known as “gravity waves”. Using the same example, if one blows too hard, the water boundary is pushed up into the air boundary, breaking the hydrogen bonds between water molecules which allows gravity to return the water to its original state. All waves in the ocean are either capillary waves or gravity waves, and their classification will be dependent on the restoring force involved.

@@

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obviously not he case. This is known as *surface elevation* and is the instantaneous elevation of the sea surface above a specific point in a time record (see figure \ref{fig:surface elevation figures}). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* or *upwardward zero-crossings* (see figure \ref{fig:surface elevation figures}). Both zero-crossings are symmetrical and essentially the same statistically. However, in practice the downward zero-crossings are preferred as it takes steepness of a approaching wave into account (the front, see figure \ref{fig:surface elevation figures}) which is relevant to characterising breaking waves. It is important to note that surface elevation can be negative while a wave profile cannot.

@@

Waves are complex and various approaches, techniques and devices have been developed in order to measure waves effectively. One of the ways that has been used extensively used in the past but less so today, are visual estimates which can be used to characterise *significant swell height* (Hs) and *significant swell period* (Tp). Visual estimates use 15 - 20 of the most well defined, higher waves of a number of wave groups to characterise Hs and Tp. Although these parameters are useful, they do not accurately reflect the waves that are occurring in nature. Ocean waves are a combination of wind sea (short, irregular, locally generated waves) and swell (long, smooth waves, generated by distant storms) and so more parameters are needed to separate Hs and Tp driven by wind sea and swell, i.e. separate Hs and Tp parameters for wind sea and swell. However, even with separate parameters for the different types of waves would still not be enough to effectively characterise waves in the ocean. In order to characterise the detail and complexity of ocean waves a different approach must be used, know as the *spectral* approach. This approach is based on the idea that the sea surface can be characterised as the summation of a large number of harmonic wave components, calculated by the *random-phase/amplitude* model. The random-phase amplitude model is a summation of wave components on a discrete scale, however a continuous scale is more relevant to characterising and measuring of waves in the ocean. The discrete spectrums are converted to continuous spectrums through various arithmetic methods. Therefore, the random-phase amplitude model is used to obtain a *continuous variance density spectrum* which is a statistical relevant measure for scientists and engineers.

@@

An object will only float if the object density ($\rho$) is less than or equal to the density of the liquid in which the object is submerged, i.e. $\rho_{object} < \rho_{liquid}$. The buoyancy force (*BF*) is the force of the displaced liquid pushing the object in an upwards direction which is equal to the weight of the displaced liquid and is expressed as: $$BF = W_{displaced-liquid}$$ The weight of the displaced liquid can be calculated by multiplying the mass of the liquid by the gravitational constant $g$ and so the weight of the displaced liquid can be calculated by: $$W_{displaced-liquid} = mg$$ The equation is expanded upon by including the calculation for mass, $m = \rho V$, and so now the weight of the displaced liquid can be calculated as: $$mg = \rho_{liquid} V g$$ where $\rho$ represents the density of the liquid and $V$ represents the portion of the object submerged. Since the object is floating, the BF can also be expressed by the weight of the object and as mentioned previously the weight of the displaced liquid can also be expressed by $mg$, which in turn can be written as $W_{displaced-liquid} = W_{object}$. Now through substituting the calculations for weight, volume and the dimensions of the object, the equation can be simplified to: $$x= h\displaystyle\frac{\rho_o}{\rho_l}$$ where $h$ is the overall height of the object, $\rho_o$ represents the density of the object and $\rho_l$ represents the density of the liquid.

It is important to note that this equation is only applicable to objects that are floating i.e. $\rho_{object} < \rho_{liquid}$. The portion of a floating object above the water surface can also be calculated using a similar approach, and instead of calculating the submerged portion ($x$) the calculation will now solve for the fraction above the surface which is represented by $y$ (see figure \ref{fig:BF_1}). The same equation for calculating density ($\rho = \displaystyle\frac{m}{V}$) will also be used in solving for $y$. As mentioned previously, for floating objects the $BF = m_{object}g$ and the BF can also be defined as the weight of the displaced liquid ($BF = W_{displaced-liquid}$). Therefore, the weight of the displaced liquid is equal to the weight of the object ($W_{displaced-liquid} = W_{object}$). On this basis equation ?? can be written as $$m_{liquid}g = m_{object}g$$ Substituting the equation for density into equation ?? the equation is now: $$\rho_{o}V_{o}g = \rho_{l}V_{l}g$$ Important to note is that $V_{o}$ represents the volume of displaced liquid i.e not the entire object volume (see figure \ref{fig:BF_1}). The constant $g$ is removed through cancellation and now the equation can be written as: $$\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$$ We also now that the volume of the object submerged is equal to the volume of the displaced liquid and the fraction $\displaystyle\frac{V_l}{V_o}$ represents the portion of the object submerged. Therefore to calculate the portion above the surface this must be subtracted by 1, therefore: $$1-\displaystyle\frac{V_l}{V_o} = y$$ and since $\displaystyle\frac{V_l}{V_o} = \displaystyle\frac{\rho_o}{\rho_l}$ the equation can now be solved via density of the liquid and the object: $$y = 1 - \displaystyle\frac{\rho_o}{\rho_l}$$ Solving the equation algebraically it can now be written as:$$ y = \displaystyle\frac{\rho_l - \rho_o}{\rho_l}$$

Once the fractions of submerged and non-submerged portions of the object have been calculated the next calculation relevant to this study would be to calculate the weight needed to cause the object to sink or as least be completely submerged. This will relate to epiphyte loading on kelp rafts in order to determine raft longevity. If a weight is placed on the object in the conceptual model that has enough weight to fully submerge (but not sink) the object then an addition buoyancy force is needed, which will be represented by $BF_2$ which is equal to the additional mass that has been added which can be expressed as $mg$. Based on the previous calculations and the laws of physics we know that: $$W_{object} = mg = BF_2 = W_{add-liq-displaced}$$ The weight of the additional liquid displaced can now be calculated by substituting the density and volume of the additional displaced liquid along with the gravitational constant $g$ which is expressed as: $$W_{AL} = \rho_{l}V_{AL}g$$
where the subscript $AL$ represents the components of *additional liquid displaced*. The $V$ can also represent the change in volume ($\Delta V$) and so the expression can be expanded to: $$W_{AL} = \rho(\Delta V) g$$} where $\Delta V$ represents the change in volume of the object, i.e $y$. To estimate the amount of input into the model the authors used a scaled approach to define the release of particles, and where based on previous research performed by @halpern2008.

@@

The effects that wind produce on floating objects is sometimes referred to as "windage factors", and is difficult to accurately and empirically describe. The difficulty is due to accurate and measurements needed for the current velocity, wind velocity, wave height and wave direction. Furthermore, these factors do not act independently and instead influence one another which compounds the complexity of calculating wind effects on drifting objects. To overcome this challenge, measurements of current and wind velocity, and wave height and direction; need to be measured along the objects trajectory for a wide range of conditions.

Wind effects or windage factors can be calculated using the approach by @griffin2017, and can be expressed as $$V = V_c + V_l + V_w$$

where $V$ is the velocity of the drifting object, $V_c$ is surface current velocity, $V_l$ is the 'leeway' velocity and $V_w$ is the velocity due to wave forces. As stated, $V_c$ is the velocity of the surface current and is averaged over the vertical extend of the drifting object. Velocity of the surface current is expressed as $$V_c = V_5 + V_s$$
where $V_5$ is the average velocity for the the top 5m of the ocean (estimated by an ocean model) and $V_s$ is the velocity effect of Stokes drift due to wave action. It is important to note that the approximation of Stokes drift does not take into account finer-scale vertical differences of the velocity field, such as Coriolis-Stokes forcing. Stokes drift generally manifests in hydrodynamic environments with short-wavelengths and locally-generated waves. @griffin2017 mitigated the exclusion of finer-scale drivers of Stokes Drift by comparing modeled surface velocities with those of *in situ* drifters.

$V_l$ is the velocity of the object relative to the water due to effects of wind force directly on the object and by expressed by $$V_l = w U_10$$

@@

$$u'= u + \frac{\delta tu_t}{2}$$ $$\upsilon'=\upsilon + \frac{\delta t\upsilon_t}{2}$$
and $u$, $\upsilon$ are orthogonal velocity components, $t$ is time, $\delta t$ is the model time step, $u_\chi$, $u_y$, $\upsilon_y$ are the $u$ and $\upsilon$ spatial velocity gradients, and $u_t$, $\upsilon_t$ are temporal gradients. Horizontal diffusion was modeled as a random walk with separate longitudinal and lateral coefficients set to simulate random turbulence. The distance increments moved by the particle at each time step is calculated as:
$$\Delta X = R_N \sqrt{6E_1\delta t}$$ $$\Delta y = R_N\sqrt{6E_2\delta t}$$
where $R_N$ is a random number in the uniform range (-1,1) and $E_1,E_2$ are the longitudinal and lateral eddy diffusivities, respectively.

@@

A study by @lebreton2012 proposed a methodology to track floating debris from source to sink based on descriptions of global waste production and ocean surface currents, using a combination of hydrodynamic and particle dispersion modelling. The particle tracking model applied uses a two stage process, (1) a hydrodynamic model solves the equations of motion to characterise the movement of water within the domain of the model, (2) virtual particles are released into the flow field established in the first step, and particles are allowed to move through hydrodynamic forcing. The authors extracted sea surface currents from the  HYCOM/NCODA ocean circulation modelling system. The HYCOM model is forced by the US Navy's Operational Global Atmospheric Prediction System (NOGAPS). The model also takes into account wind stress, wind speed, heat flux, and precipitation; and was looped five times to accurately represent ocean circulation patterns for 30 years. The authors used the velocity data aquired from HYCOM and coupled it to the Lagrangian particle tracking model Pol3DD which was used to simulate dispersion. No additional wind stress terms were applied to the motion of particles, as this factor was already expressed in the HYCOM hydrodynamic data. Potential extra stress for emerged parts of debris was considered negligible as all floating debris are assumed to be fully submerged. The Pol3DD model was also modified so provide additional information on individual particles such as origin, age, and trajectory information of individual particles. Since the model considered a spherical earth, particles were allowed to re-enter the simulation if the passed around the 'globe'. In this study only the surface layer was considered as the principle layer which determines the fate of floating particles i.e anthorpogenic debris.
